
\section{PCA Extensions}

\label{chap:extensions}

In addition to the core component framework, \pcasp includes implementations
for a variety of components that support commonly used functionality. 
These extensions of \pcasp are available in PyUtilib packages separate 
from the \pcasp core.  This emphasizes the modularity of the \pca, 
and it illustrates how to define \pcasp components that are 
automatically registered as part of an application.
The following sections briefly describe these \pcasp extensions.


\subsection{Component Loaders}

\label{sec:loaders}

\pcasp components can be loaded from either Python modules or Python eggs. This
capability supports the runtime extension of the \pca,
which has proven very powerful in frameworks like Trac. 
Component services for loading are defined in the \code{pyutilib.component.loader} package.
The core \pcasp
defines extension points that use these services, which can be
applied as follows:
\begin{quotation}
\begin{lstlisting}
import sys
import os
env = os.environ["PATH"]
PluginGlobals.load_services(path=env.split(os.sep))
\end{lstlisting}
\end{quotation}
In this example, the user's \code{PATH} environment is used to define the list
of directories that are searched for Python modules and eggs.  

The \code{load\_services} takes two other optional arguments that control
how components are loaded.  The \code{name\_re} argument can be used to
define a regular expression that filters the files in the directories
that are searched.  The following shows how to specify that services starting with \textit{my}
are loaded:
\begin{quotation}
\begin{lstlisting}
PluginGlobals.load_services(name_re="my.*")
\end{lstlisting}
\end{quotation}

By default, when services are loaded they are disabled.  This facilitates the management of
services in complex applications using 
configuration files (see below).  The \code{auto\_disable} flag can be used
to automatically activate services:
\begin{quotation}
\begin{lstlisting}
PluginGlobals.load_services(auto_disable=False)
\end{lstlisting}
\end{quotation}


\subsection{Registering Executables}

The \code{pyutilib.component.executable} package defines the \code{ExternalExecutable} plugin, which is used to define services that provide
information about external executables. Services declare the executable
name and user documentation, and then service methods indicate whether
the executable is enabled (i.e. whether it is found, and the path of
the executable:
\begin{quotation}
\begin{lstlisting}
service = ExternalExecutable(name='ls', 
            doc='A utility to list file in Unix operating systems')

service.enabled()     
# Returns True if the executable is found on the user path.

service.get_path()
# Returns a string that defines the path to this executable, 
# or None if service is disabled.
\end{lstlisting}
\end{quotation}
The registration process is simplified with the \code{pyutilib.services} package, 
which includes the \code{register\_executable} function:
\begin{quotation}
\begin{lstlisting}
import pyutilib.services

pyutilib.services.register_executable('zip')
\end{lstlisting}
\end{quotation}
This function searches the user's \code{PATH} environment for the \code{zip} executable (or \code{zip.exe} on Windows machines).  

A developer can use the \code{registered\_executable} function to the
access the absolute path of a registered executable.  If the executable
is not found in the user's \code{PATH}, then this returns \code{None}.
Also, if no executable is specified, then this function returns a list
of all registered executables.


\subsection{Temporary Files}

The \code{pyutilib.component.config} packages provides a services for managing temporary files.  The \code{TempfileManager} object is a component service whose methods can be used to create and
cleanup temporary files;  for convenience, this object is accessible from the \code{pyutilib.services} package.

The main method in this service is \code{create\_tempfile}, which can create a temporary file with a specified suffix and prefix in a specified directory:
\begin{quotation}
\begin{lstlisting}
import pyutilib.services

pyutilib.services.create_tempfile(prefix='myfile',
                                  suffix='.txt', dir='/home/jdoe')
\end{lstlisting}
\end{quotation}
By default, this service creates unique filenames.  However, if the \code{sequential\_files}
method is called, then the body of the temporary files will be an integer that 
is incremented every time a temporary file is created.  Although these filenames may not
be unique, this sequential naming scheme may make it easier to diagnose errors in a
complex application.

This service keeps track of the temporary files that it creates.  This allows an application
developer to avoid this bookkeeping, and instead rely on this service to delete temporary 
files with the \code{clear\_tempfiles} method.  Furthermore, a developer can explicitly
declare a file as temporary using the \code{add\_tempfile} method, thereby allowing this
service to delete it.


\subsection{Options and Configuration Files}

The \code{pyutilib.component.config} package defines interfaces and plugins
for managing service options. The \code{Configuration} service is used to manage
the global configuration of all services. This class coordinates with
\code{Option} services. Plugins can declare options with the \code{declare\_option}
method, which registers these options with the \code{Configuration} service. This
service reads and writes options to configuration files (using Python's
\code{ConfigParser} package).

This package also declares the \code{ManagedPlugin} and \code{ManagedSingletonPlugin}
classes, which are plugin base classes that include options that can be
used to enable or disable services using the \code{Configuration} service. 


\subsubsection{Configuration Files}

A \pcasp configuration file consists of a list of sections.  Each section is lead by a \code{[section]} header, and a section contains a list of \code{name = value} entries.  For example, 
the following configuration file consists of two sections with four option values:
\begin{quotation}
\begin{lstlisting}
# COMMENT
[globals]
a = 1
b = /dev/null
c = 1,2,3
[a.b]
zz = 4.5
\end{lstlisting}
\end{quotation}
\pcasp plugin classes declare component options with the \code{declare\_option} method, which is defined in
\code{pyutilib.component.doc}.
These options can be initialized with a configuration file.  For example, the following plugin
declares four options.  
\begin{quotation}
\begin{lstlisting}
class PluginWithOptions(Plugin):
    def __init__(self):
        declare_option("a")
        declare_option("b")
        declare_option("c")
        declare_option("zz",section='a.b')
\end{lstlisting}
\end{quotation}
The default configuration section for an option is \code{[globals]},
but the option declaration can specification the section name.  For example, the configuration file
described above can be used to initialize the \code{PluginWithOptions} plugin.

The \code{Configuration} class manages loading and storing configuration data for the options
that are registered by \pcasp services.  This class defines the following methods:
\begin{description}

\item[clear] Clear the configuration data.
\item[load] Load configuration data from a file.
\item[pprint] Print a simple summary of the configuration data.
\item[save] Write configuration data to a file.
\item[sections] Return a list of the sections that have been loaded.
\item[summarize] Summarize the options that have been registered with the \pca.

\end{description}
Once data is loaded with the \code{load} method, the \code{sections} method can be used to provide a list of the sections that were loaded.  The Python \code{\_\_contains\_\_} method can 
also be used to check if a section was loaded:
\begin{quotation}
\begin{lstlisting}
from pyutilib.component.config import *

config = Configuration()

# Load configuration data
config.load('config.ini')

# Check if the 'globals' section was loaded
if 'globals' in config:
    print "The 'globals' section was loaded"

# Get the 'globals' section
section = config['globals']
\end{lstlisting}
\end{quotation}
The Python \code{\_\_getitem\_\_} method is used at the end of this example to get the data for the 'globals' section.  Section data consists of a dictionary that maps option names to value strings.  Note that the \code{Configuration} class automatically loads this data into the 
corresponding options that have been registered with the \pca.


\subsubsection{Declaring Options}

The \code{declare\_option} creates an \code{Option} object that is a data member in a 
plugin.  The standard syntax for this function is to specify the option name, which is used to define an attribute in the plugin with the same name:
\begin{quotation}
\begin{lstlisting}
class TmpPlugin(SingletonPlugin):

  declare_option('x')
\end{lstlisting}
\end{quotation}
The \code{local\_name} keyword can be used to specify a different name for this attribute
within the plugin.  For example, consider:
\begin{quotation}
\begin{lstlisting}
class TmpPlugin(SingletonPlugin):

  declare_option('x', local_name='y')
\end{lstlisting}
\end{quotation}
In this example, the option is declared with name \code{x} in the \pcasp
registry, but it has attribute name \code{y} within this plugin.  
The \code{default}
keyword defines the default value of an option, and the \code{doc} option specifies a
document string that describes the option;  this information is used by the \code{Configuration} class when printing option summary information.

As noted earlier, the \code{section} keyword can be used to specify
the section in configuration data that this option is expected.
The \code{section\_re} keyword supports a more generic mechanism.
If the \code{section\_re} is specified with a regular expression,
then this option will be initialized from any section that matches this
regular expression.  If sections match and contain data for this option,
then the last section specified in the configuration data will be used
to initialize this option.

Finally, the \code{cls} keyword specifies the option type.  Option types are described in 
the next section.

% TODO: should we change the syntax to 'type' instead of 'cls'?  I think that 'type' might
% be a reserved word...


\subsubsection{Options Types}

The default option type is \code{Option}.  These options treat option values as strings, 
even when they could be interpreted as numeric values.  The \code{BoolOption}, \code{IntOption} and \code{FloatOption} types respectively interpret option values as booleans, integers and floating point values.  The \code{OptionError} exception is raised if the option value is not the appropriate value.

The \code{FileOption} type interprets the option value into a path.
A relative path is converted to an absolute path using the path for
the configuration file.  Thus, a user can load file path data from
any directory but specify the file data relative to the path of the
application configuration.  In addition, this option type also supports
a \code{directory} keyword that can be used to specify how relative
paths are resolved.  The \code{ExecutableOption} type is an extension of
\code{FileOption} that confirms that the file can be executed.  If the
file name does not include path information, then \pcasp will search
for the executable using the user's \code{PATH} environment before
initializing this option.

The \code{DictOption} type supports an interface to all options in a section.
For example, consider:
\begin{quotation}
\begin{lstlisting}
class TmpPlugin(Plugin):

    options = DictOption(section="bar")
\end{lstlisting}
\end{quotation}
The \code{options} object will be populated by all data in the \code{bar} section.
For example, if the names \code{a} and \code{b} are defined in this section, then
they can be referenced as \code{options.a} and \code{options.b}.  Similarly,
data can be inserted into the \code{bar} section by simply 
specifying the value of attributes of this object:
\begin{quotation}
\begin{lstlisting}
options.c = 2
\end{lstlisting}
\end{quotation}


\subsubsection{Using Options in Services}

It is convenient for singleton plugins to declare options as part of the class definition:
\begin{quotation}
\begin{lstlisting}
class Plugin1(SingletonPlugin):

  declare_option("x")
\end{lstlisting}
\end{quotation}
This type of declaration makes sense since there is a single instance of the class \code{Plugin1}.
For non-singleton plugins, this type of declaration would make the \textit{same} option data
available to all instances of the plugin.  To declare different options for different 
non-singleton plugin instances, it suffices to execute \code{declare\_option} within the
plugin constructor:
\begin{quotation}
\begin{lstlisting}
class Plugin2(Plugin):

    def __init__(self, sec):
        declare_option("x", section=sec)
\end{lstlisting}
\end{quotation}
Note that this example allows the \code{Plugin2} services to distinguish the configuration of
these different \code{x} options by specifying different sections in the configuration file.
Although this is not required, this is often desirable in practice.


\subsubsection{Managed Services}

The \pcasp supports explicit management of services using the configuration management
technology described in this section.  The \code{ManagedPlugin} and \code{ManagedSingletonPlugin}
classes include an \code{enable} option that controls whether their corresponding
services are activated.  When services are loaded from EGG files and modules, they
are disabled by default.  The \code{Services} section of a configuration file can
be used to activate these plugins:
\begin{quotation}
\begin{lstlisting}
[Services]
plugin1 = True
plugin3 = True
\end{lstlisting}
\end{quotation}
This allows an application administrator to install a variety number of application
services that a user selectively enables.


\subsection{Other Extensions}

The following extension packages include plugins and applications interfaces for \pcasp applications.  These extension packages are less mature, and consequently they are not documented in detail right now.

\paragraph{pyutilib.component.config}  This extension package defines a plugin that manages logging of \pcasp actions.

\paragraph{pyutilib.component.app}  This package defines a simple application class that 
can be used as the basis of a component-based application.  This application class provides support for managing configuration from a configuration file, and for managing logging activity.




